\textit{Supporting Information for}: Memory for television episodes preserves event content while introducing new across-event similarities

### 2020-01-08 paper/supporting_information.tex
finished embedding supporting info section

\begin{figure}[p!]
\centering
\includegraphics[width=1\textwidth]{figs/parameter_search}
\caption{\small \textbf{Optimizing topic model parameters.}  We performed a grid search over video sliding window length ($\omega \in \{5, 10, 25, 50, 100 \}$), recall sliding window length ($\rho \in \{5, 10, 25, 50, 100 \}$), and number of topics ($K \in \{5, 10, 25, 50, 100 \}$).  The reported correlations are between per-subject video-recall trajectory correlations and per-subject hand-annotated memory performance ratings.}
\label{fig:paramsearch}
\end{figure}

@@

\begin{figure}[p!]
\centering
\includegraphics[width=0.65\textwidth]{figs/topic_words}
\caption{\small \textbf{Topics discovered in \textit{Sherlock}.} We applied a topic model to hand-annotated information about 1000 scenes spanning the 45 minute episode.  We identified 27 unique topics with non-zero weights (we used $K=100$ topics to fit the model).  Each topic comprises a distribution of weights over all words in the vocabulary.  For each topic, we show the words with the 10 largest weights, along with a suggested description of the topic.}
\label{fig:topics}
\end{figure}
\FloatBarrier

\subsection*{Feature importance analyses}
To determine the contribution of each feature to the structure of the video topic proportions, we conducted a ``leave one out'' analysis.  Specifically, we compared the original video topic trajectory (created using all hand-annotated features from the 1000 hand-annotated scenes spanning the \textit{Sherlock} episode; see \textit{Methods} for a full list of features) with video trajectories created using all but one type of feature.  We created temporal correlation matrices for each trajectory (using the topic proportions matrices) and correlated the upper triangles of each impoverished trajectory with the original feature-complete trajectory.  Observing a lower correlation between an impoverished trajectory (holding out a particular feature) and the feature-complete trajectory would suggest that the given feature played a more prominent role in shaping the temporal structure of the feature-complete trajectory.  We found that hand-annotated scene locations provided the most structure to the feature-complete trajectory, whereas the name of the character(s) in focus for each shot provided the least structure (Fig.~\ref{fig:feature-importance}A).

\begin{figure}[t]
\centering
\includegraphics[width=1\textwidth]{figs/feature_value}
\caption{\small \textbf{Feature importance analysis.} \textbf{A.} Contributions of each feature type to the structure of the video trajectory. The bar heights reflect the correlation between the video trajectory computed using all features with a video trajectory computed using all features except the indicated feature.  (Lower bars reflect features that contribute more substantially to the video trajectory's shape.) \textbf{B.} Which features are preserved during recall?  The bar heights reflect the (average) across-participant correlations between the video and recall trajectories.  Error bars denote bootstrap-estimated standard error of the mean.  \textbf{C.} Feature correlation matrix.  Each entry displays the correlation between video topic trajectories created using only the indicated (row/column) features.}
\label{fig:feature-importance}
\end{figure}

@@

Next, we wondered how the different types of features might relate.  For example, knowing which character is in focus during a given scene may also provide information about which character is speaking.  We computed video topic trajectories for each feature in turn, and then compared the upper triangles of the temporal correlation matrices of all pairs of features.  This provided additional confirmation that the shape of the full trajectory (including all types of features) was largely driven by narrative details.  We also found that character-driven features (characters on screen, characters speaking, and characters in focus) were strongly correlated.  Other details, such as the presence or absence of music, led to very different topic trajectories (Fig.~\ref{fig:feature-importance}C).
\FloatBarrier

@@

Two other parameters ultimately affect this balance between preserving local and global structure: the seed ($\tau$) for the (pseudo-)random number generator, and the order of the observations (i.e., trajectories) to be embedded ($\phi$).  As described in \textit{Methods}, we ensured the video and recall events were projected onto the \textit{same} low-dimensional manifold by fitting the embedding model to a stacked matrix of all video, average recall, and individuals' recall events.  After initializing the low-dimensional simplicial set (by default, using a spectral embedding of the high-dimensional simplicial set's fuzzy 1-skeleton), UMAP optimizes the embedding using stochastic gradient descent with cross-entropy as a cost function.  During optimization, indices of the data are sampled at random, and thus local minimum reached by the optimization is dependent on both the state of the RNG and the sequence of observations.

We performed a grid search over pre-specified values of these hyper-parameters, and optimized the manifold space for visualization based on two criteria: First, that the 2D embedding of the video trajectory should reflect its original 100-dimensional structure as faithfully as possible. Second, that the path traversed by the video trajectory should intersect itself as few times as possible.  The first criteria helps promote the ability to gain (albeit fuzzy) intuitions about similarities and differences between sections of video content based on their locations in the embedding space.  The second criteria was motivated by the low off-diagonal values in the video trajectory's temporal correlation matrix (suggesting that the same topic-space coordinates should not be revisited; see Figure~2A in the main text) and the inability to accurately describe the consistency of participants' transitions through topic space at locations where those transitions were consistent, but occurred in multiple, orthogonal directions (see \textit{Methods} for analysis details).  We formalized this optimization as the combination of values that satisfied
\[
\argmax_{\left\{K, \gamma, \delta, \tau, \varphi~\in~E~\mathrm{|}~\Phi\left(K, \gamma, \delta, \tau, \varphi \right) \right\}} \left[\mathrm{corr}\left(\Upsilon\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right),~\Upsilon\left(\chi\right)\right)\right],~\mathrm{where}
\]
\[
\Phi\left(K, \gamma, \delta, \tau, \varphi \right) = \argmin_{K, \gamma, \delta, \tau, \varphi} \left[\Gamma\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right)\right],
\]
$\xi$ is the video's trajectory through the manifold space; $\chi$ is the original, 100-dimensional video trajectory; $\Upsilon$ is a condensed matrix of pairwise distances between event vectors (computed using correlation distance in the original topic space and Euclidean distance in the manifold space); $\mathrm{corr}\left(\Upsilon\left(\xi\right), \Upsilon\left(\chi\right)\right)$ is the correlation between the sets of pairwise distances, and $\Gamma$ is the number of instances in which lines drawn between consecutive video event embeddings intersect.   We searched over the hyper-parameter values
\[
K \in \{

\section*{Participant-level figures referenced in the main text}

@@

\

>>>>

\begin{figure}[]
\centering
\includegraphics[width=1\textwidth]{figs/parameter_search}
\caption{\small \textbf{Optimizing topic model parameters.}  We performed a grid search over video sliding window length ($\omega \in \{5, 10, 25, 50, 100 \}$), recall sliding window length ($\rho \in \{5, 10, 25, 50, 100 \}$), and number of topics ($K \in \{5, 10, 25, 50, 100 \}$).  The reported correlations are between per-subject video-recall trajectory correlations and per-subject hand-annotated memory performance ratings.}
\label{fig:paramsearch}
\end{figure}

@@

\begin{figure}[]
\centering
\includegraphics[width=0.65\textwidth]{figs/topic_words}
\caption{\small \textbf{Topics discovered in \textit{Sherlock}.} We applied a topic model to hand-annotated information about 1000 scenes spanning the 45 minute episode.  We identified 27 unique topics with non-zero weights (we used $K=100$ topics to fit the model).  Each topic comprises a distribution of weights over all words in the vocabulary.  For each topic, we show the words with the 10 largest weights, along with a suggested description of the topic.}
\label{fig:topics}
\end{figure}
%\FloatBarrier

\subsection*{Feature importance analyses}
To determine the contribution of each feature to the structure of the video topic proportions, we conducted a ``leave one out'' analysis.  Specifically, we compared the original video topic trajectory (created using all hand-annotated features from the 998 hand-annotated scenes spanning the \textit{Sherlock} episode; see \textit{Methods} for a full list of features) with video trajectories created using all but one type of feature.  We created temporal correlation matrices for each trajectory (using the topic proportions matrices) and correlated the upper triangles of each impoverished trajectory with the original feature-complete trajectory.  Observing a lower correlation between an impoverished trajectory (holding out a particular feature) and the feature-complete trajectory would suggest that the given feature played a more prominent role in shaping the temporal structure of the feature-complete trajectory.  We found that hand-annotated scene locations provided the most structure to the feature-complete trajectory, whereas the name of the character(s) in focus for each shot provided the least structure (Fig.~\ref{fig:feature-importance}A).

\begin{figure}[]
\centering
\includegraphics[width=1\textwidth]{figs/feature_value}
\caption{\small \textbf{Feature importance analysis.} \textbf{A.} Contributions of each feature type to the structure of the video trajectory. The bar heights reflect the correlation between the video trajectory computed using all features with a video trajectory computed using all features except the indicated feature.  (Lower bars reflect features that contribute more substantially to the video trajectory's shape.) \textbf{B.} Which features are preserved during recall?  The bar heights reflect the (average) across-participant correlations between the video and recall trajectories.  Error bars denote bootstrap-estimated standard error of the mean.  \textbf{C.} Feature correlation matrix.  Each entry displays the correlation between video topic trajectories created using only the indicated (row/column) features.}
\label{fig:feature-importance}
\end{figure}

@@

Next, we wondered how the different types of features might relate.  For example, knowing which character is in focus during a given scene may also provide information about which character is speaking.  We computed video topic trajectories for each feature in turn, and then compared the upper triangles of the temporal correlation matrices of all pairs of features.  This provided additional confirmation that the shape of the full trajectory (including all types of features) was largely driven by narrative details.  We also found that character-driven features (characters on screen, characters speaking, and characters in focus) were strongly correlated.  Other details, such as the presence or absence of music, led to very different topic trajectories (Fig.~\ref{fig:feature-importance}C).
%\FloatBarrier

@@

Two other parameters ultimately affect this balance between preserving local and global structure: the seed ($\tau$) for the (pseudo-)random number generator, and the order of the observations (i.e., trajectories) to be embedded ($\phi$).  As described in \textit{Methods}, we ensured the video and recall events were projected onto the \textit{same} low-dimensional manifold by fitting the embedding model to a stacked matrix of all video, average recall, and individuals' recall events.  After initializing the low-dimensional simplicial set (by default, using a spectral embedding of the high-dimensional simplicial set's fuzzy 1-skeleton), UMAP optimizes the embedding using stochastic gradient descent with cross-entropy as a cost function.  During optimization, indices of the data are sampled at random, and thus local minimum reached by the optimization is dependent on both the state of the RNG and the sequence in which observations are concatenated.

We performed a grid search over pre-specified values of these hyper-parameters, and optimized the manifold space for visualization based on two criteria: First, that the 2D embedding of the video trajectory should reflect its original 100-dimensional structure as faithfully as possible. Second, that the path traversed by the video trajectory should intersect itself as few times as possible.  The first criteria helps promote the ability to gain (albeit fuzzy) intuitions about similarities and differences between sections of video content based on their locations in the embedding space.  The second criteria was motivated by the low off-diagonal values in the video trajectory's temporal correlation matrix (suggesting that the same topic-space coordinates should not be revisited; see Figure~2A in the main text) and the inability to accurately describe the consistency of participants' transitions through topic space at locations where those transitions were consistent, but occurred in multiple, orthogonal directions (see \textit{Methods} for analysis details).  We formalized this optimization as the combination of values that satisfied
\[
\argmax_{\left\{K, \gamma, \delta, \tau, \varphi~\in~E~\mathrm{|}~\Phi\left(K, \gamma, \delta, \tau, \varphi \right) \right\}} \left[\mathrm{corr}\left(\Upsilon\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right),~\Upsilon\left(\chi\right)\right)\right],~\mathrm{where}
\]
\[
\Phi\left(K, \gamma, \delta, \tau, \varphi \right) = \argmin_{K, \gamma, \delta, \tau, \varphi} \left[\Gamma\left(\xi_{K, \gamma, \delta, \tau, \varphi}\right)\right],
\]
$\xi$ is the video's trajectory through the manifold space; $\chi$ is the original, 100-dimensional video trajectory; $\Upsilon$ is a condensed matrix of pairwise distances between event vectors (computed using correlation distance in the original topic space and Euclidean distance in the manifold space); $\mathrm{corr}\left(\Upsilon\left(\xi\right), \Upsilon\left(\chi\right)\right)$ is the correlation between the sets of pairwise distances, and $\Gamma$ is the number of instances in which lines drawn between consecutive video event embeddings intersect.   The sets of hyper-parameter values we searched over comprised: $K \in \{10x \in \mathbb{Z}~|~10 < x <  22\} \cup \{161\}$ (a range roughly centered on half the total number of events, 161, in order to balance representations of local and global structure); $\gamma \in \{1, 3, 5, 7, 9\}$; $\delta \in \{0.1, 0.3, 0.5, 0.7\}$; $\tau \in \{x \in \mathbb{Z}~|~0 < x < 101\}$; and $\phi \in {S\choose3},~\mathrm{where}~S=\{\mathrm{video~events,~average~recall~events, individual~recall~events}\}$.  The optimal parameters (that yielded $\Phi=0$) were $K=170$, $\gamma=7$, $\delta=0.7$, $\tau=41$, and $\varphi$ as a the sequences of average recall events, video events, and individuals' recall events, vertically concatenated.

\FloatBarrier
\section*{Participant-level figures referenced in the main text}

@@

\FloatBarrier
